---
title: "Chapter 5 notes"
author: "Bryce Watson"
date: "10/21/2024"
output:
  html_document:
    theme:
      bootswatch: cerulean
---

### Review of R for Data Science by Hadley Wickam - Ch. 1

### Begin ggplot development
```{r}
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
```

ggplot(data = penguins) # Unnecessary as separate step.
head(penguins)

# Load data
```{r, message=FALSE}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = species)
) + 
  geom_point() +
  geom_smooth(method = "lm")
```

It’s generally not a good idea to represent information using only colors on a plot, as people perceive colors differently due to color blindness or other color vision differences. Therefore, in addition to color, we can also map species to the shape aesthetic.

```{r, message=FALSE}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) + 
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = "lm")

```