---
title: "Chapter 5 notes"
author: "Bryce Watson"
date: "10/21/2024"
output:
  html_document:
    theme:
      bootswatch: cerulean
---

# Review of R for Data Science by Hadley Wickam - Ch. 1

### Begin ggplot development
```{r}
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
```

ggplot(data = penguins) # Unnecessary as separate step.
head(penguins)

### Load data
```{r, message=FALSE, warning=FALSE}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = species)
) + 
  geom_point() +
  geom_smooth(method = "lm")
```

It’s generally not a good idea to represent information using only colors on a plot, as people perceive colors differently due to color blindness or other color vision differences. Therefore, in addition to color, we can also map species to the shape aesthetic.

```{r, message=FALSE, warning=FALSE}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) + 
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = "lm")

```

Final plot - Added colorblind color scale, title/subtitle, lables

```{r, message=FALSE, warning=FALSE}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) + 
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = "lm") +
  labs(
    title = "Body mass and flipper length",
    subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
    x = "Flipper length (mm)", y = "Body mass (g)",
    color = "Species", shape = "Species"
  ) +
  scale_color_colorblind()

```

### Excercise 8 from R for Data Science (2e), 1.2.5
Recreate the following visualization. What aesthetic should bill_depth_mm be mapped to? And should it be mapped at the global level or at the geom level?

Below are my original edits (I used chatGPT to look up the difference between "gam" and "loess" methods and went with loess since the dataset is relatively small).
decided the bill_depth_mm should be mapped as an aes(color) modifier at the geom level. 

```{r, message=FALSE, warning=FALSE}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) + 
  geom_point(mapping = aes(color = bill_depth_mm)) +
  geom_smooth(method = "loess")


```

# Review of Data Visualization by Kieran Healy - Ch 3.

Install packages from book.
```{r}
install.packages(c("tidyverse", "broom", "coefplot", "cowplot",
                 "gapminder", "GGally", "ggrepel", "ggridges", "gridExtra",
                 "here", "interplot", "margins", "maps", "mapproj",
                 "mapdata", "MASS", "quantreg", "rlang", "scales",
                 "survey", "srvyr", "viridis", "viridisLite", "devtools"))
devtools::install_github("kjhealy/socviz")
library()
```

Book not loading, try again later. 

# Telling Stories with Data - Ch. 5

### Libraries used in this chapter.
library(carData)
library(datasauRus)
library(ggmap)
library(janitor)
library(knitr)
library(maps)
library(mapproj)
library(modelsummary)
library(opendatatoronto)
library(patchwork)
library(tidygeocoder)
library(tidyverse)
library(troopdata)
library(WDI)

### 5.2 Graphs - Based on: https://juliasilge.com/blog/datasaurus-multiclass/
```{r, message=FALSe}
datasaurus_dozen |>
  filter(dataset %in% c("dino", "star", "away", "bullseye")) |>
  summarise(across(c(x, y), list(mean = mean, sd = sd, median = median)),
            .by = dataset) |>
  kable(col.names = c("Dataset", "x mean", "x median", "x sd", "y mean", "y sd", "y median"),
        booktabs = TRUE, digits = 1)
```

### 5.2.1 Bar Charts   
```{r, message=FALSE}
beps <- 
  BEPS |>
  as_tibble() |>
  clean_names() |>
  select(age, vote, gender, political_knowledge)

head(beps)
```

### 5.4 Maps
```{r, results='hide'}
france <- map_data(map = "france")

head(france)

french_cities <-
  world.cities |>
  filter(country.etc == "France")

head(french_cities)
```

Plot the gathered map data.

```{r, message=FALSE, warning=FALSE}
ggplot() +
  geom_polygon(
    data = france,
    aes(x = long, y = lat, group = group),
    fill = "white",
    colour = "grey"
  ) +
  coord_map() +
  geom_point(
    aes(x = french_cities$long, y = french_cities$lat),
    alpha = 0.3,
    color = "royalblue",
    size = 0.5
  ) +
  theme_dark() +
  labs(x = "Longitude", y = "Latitude")

```

### Australian polling places

Using ggmap we need to define a bounding box with two lattitudes and two longitudes.
```{r}
bbox <- c(left = 148.95, bottom = -35.5, right = 149.3, top = -35.1)

```

Register account: https://client.stadiamaps.com/signup/ 

```{r, message=FALSE, error=FALSE}
register_stadiamaps(key = "bc0f708a-97be-4d87-b5cf-a0155debf46b", write = TRUE)
get_stadiamap()

canberra_stamen_map <- get_stadiamap(bbox, zoom = 11, maptype = "stamen_toner_lite")

ggmap(canberra_stamen_map)

```


Plot voting booth data on top of tile.
```{r, message=FALSE, error=FALSE}
# Download data:

booths <-
  read_csv(
    "https://results.aec.gov.au/24310/Website/Downloads/GeneralPollingPlacesDownload-24310.csv",
    skip = 1,
    guess_max = 10000
  )

# Filter for booths close to Canberra

booths_reduced <-
  booths |>
  filter(State == "ACT") |>
  select(PollingPlaceID, DivisionNm, Latitude, Longitude) |>
  filter(!is.na(Longitude)) |> # Remove rows without geography
  filter(Longitude < 165) # Remove Norfolk Island

# Plot booths to underlying tile.

ggmap(canberra_stamen_map, extent = "normal", maprange = FALSE) +
  geom_point(data = booths_reduced,
             aes(x = Longitude, y = Latitude, colour = DivisionNm),
             alpha = 0.7) +
  scale_color_brewer(name = "2019 Division", palette = "Set1") +
  coord_map(
    projection = "mercator",
    xlim = c(attr(map, "bb")$ll.lon, attr(map, "bb")$ur.lon),
    ylim = c(attr(map, "bb")$ll.lat, attr(map, "bb")$ur.lat)
  ) +
  labs(x = "Longitude",
       y = "Latitude") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

### 5.4.3 Geocoding

Using tidygeocoder
```{r, message=FALSE, warning=FALSE}
place_names <-
  tibble(
    city = c("Sydney", "Toronto", "Accra", "Guayaquil"),
    country = c("Australia", "Canada", "Ghana", "Ecuador")
  )

# geocode with function geo()
place_names <-
  geo(
    city = place_names$city,
    country = place_names$country,
    method = "osm"
  )

# plot data on map
world <- map_data(map = "world")

ggplot() +
  geom_polygon(
    data = world,
    aes(x = long, y = lat, group = group),
    fill = "white",
    colour = "grey"
  ) +
  geom_point(
    aes(x = place_names$long, y = place_names$lat),
    color = "royalblue") +
  geom_text(
    aes(x = place_names$long, y = place_names$lat, label = place_names$city),
    nudge_y = -5) +
  theme_minimal() +
  labs(x = "Longitude",
       y = "Latitude")
```
