---
title: "Chapter 3 notes"
author: "Bryce Watson"
date: "10/20/2024"
output:
  html_document:
    theme:
      bootswatch: cerulean
---

### 3.5.5 Tests
 1. **Boundary conditions** - testing extreme/edge cases (min/ma inputs)
 2. **Classes** - test variable data types
 3. **Missing data** - check how code handles missing or null values
 4. **The number of observations and variables** - verify the correct # of observations and variables are being processed.
 5. **Duplicates** - test for duplicate data.
 6. **Regression results** - ensure regression outputs are not unexpected after code changes.

Run at least each of these relevant tests during the simulation phase. Other tests may be needed. 


```{r}
fu_name <- c(1:5)
```

### 3.6.3 Code review ###

Minimal steps to keep in mind during code review:

 - Is there an informative README and how could it be improved?
 - Are the file names and variable names consistent, informative, and meaningful?
 - Do the comments allow you to understand why something is being done?
 - Are the tests both appropriate and sufficient? Are there edge cases or corner solutions that are not considered?  
 - Similarly, are there unnecessary tests that could be removed?
 - Are there magic numbers that could be changed to variables and explained?
 - Is there duplicated code that could be changed?
 - Are there any outstanding warnings that should be addressed?
 - Are there any especially large functions or pipes that could be separated into smaller ones?
 - Is the structure of the project appropriate?
 - Can we change any of the code to data (Irving et al. 2021, 462)?
 
Ensure code is readable by at least one person. Preferably it is reviewed by a team. Limit code reviews to only a few hundred lines and around one hour so as to maintain efficiency.

### 3.8 Excercises ###

(Plan) Consider the following scenario: In a certain country there are only ever four parties that could win a seat in parliament. Whichever candidate has a plurality of votes in the area associated with a given seat wins that seat. The parliament is made up of 175 total seats. An analyst is interested in the number of votes for each party by seat. Please sketch what a dataset could look like, and then sketch a graph that you could build to show all observations.

```{r}
# Create a dataset for 175 seats
n_seats <- 175

# Define parties and set up seat numbers
parties <- c("Republican", "Democrat", "Libertarian", "Green")
seat_numbers <- rep(1:n_seats, each = 4)

# Generate random vote counts (adjust range as needed)
set.seed(123)
votes <- sample(500:15000, size = n_seats * 4, replace = TRUE)

# Create data frame
election_data <- data.frame(
  Candidate = paste("Candidate", rep(1:(n_seats *4))),
  Party_Affiliation = rep(parties, times = n_seats),
  Vote_Count = votes,
  Seat_Number = seat_numbers
)

# View data
head(election_data)

```

```{r, message=FALSE}
# Create graph showing vote results by seat.
library(ggplot2)
library(magrittr)
library(tidyverse)

seat_holders <- election_data %>%
  group_by(Seat_Number) %>%
  filter(Vote_Count == max(Vote_Count)) %>%
  ungroup()

# Assign colors to parties
party_colors <- c("Republican" = "red", "Democrat" = "blue",
  "Libertarian" = "yellow", "Green" = "green")

# Update the party affiliation to match the winning parties

seat_holders$Party_Affiliation <- factor(seat_holders$Party_Affiliation,
                                         levels = c("Republican", "Democrat",                                            "Libertarian", "Green"))

# Create bar graph
ggplot(seat_holders, aes(x = Party_Affiliation, fill = Party_Affiliation)) +
  geom_bar() +
  scale_fill_manual(values = party_colors) + 
  labs(title = "Seats won by each party", x = "Party", y = "Number of seats") +
  theme_minimal()
```
